|

### 2021-10-05 03-Chp-Systematic-Review.Rmd
End of day:

# Systematic review of all evidence available on the association between blood lipids (and treatment) and dementia outcomes {#sys-rev-heading}

@@

<!-- TODO Extra studies to include: -->
<!-- * Larsson - 10.1136/bmj.j5375 [@larsson2017b] -->
<!-- * Statins and the risk of dementia Jick -->

@@

Systematic reviews are a type of research that aim to use all existing evidence to provide the best answer to an important research question. They do this by finding and combining the results from many related primary research studies. Reviews involve multiple steps including: searching of existing studies; assessment of the studies against predefined inclusion criteria; collection of data from each study; assessment of each study's methods.

This chapter presents a systematic review of primary studies that have examined the relationship between the levels of blood lipids (such as cholesterol and triglycerides), and treatments that change these levels, and dementia.

There were `n_included` primary studies that contained information on this relationship. I found that statins reduce the risk of Alzheimer's disease, but had no effect of vascular dementia. Lipids were not associated with any outcome. The methods used in some of the primary studies meant that I was less confident in the accuracy of their results.

@@

In this chapter, I describe a comprehensive systematic review of the relationship between blood lipid levels, and treatments that modify them, and the subsequent risk of dementia and related outcomes. This analysis sought to address two specific aims.

Firstly, as discussed in the Introduction to this thesis (Section \@ref(evidence-association)), several diverse forms of evidence on the relationship of lipids and dementia exist. These include randomised controlled trials, observational studies of different analytical design, and Mendelian randomisation studies. However, based on a scoping review of existing literature, no previous evidence synthesis exercise has attempted to examine the association of lipids/statins with dementia outcomes across these distinct evidence types. Collating these diverse evidence sources is important, as if the observed association between lipids and dementia is constant across them, it increases our confidence in the association. As such, the primary aim of this analysis was to systematically review all available literature, regardless of study design.

Secondly, I explicitly sought to include health-related preprint servers as a potential evidence source in this review, as they are infrequently considered by evidence synthesists but may contain important unpublished studies. As a sensitivity analysis to the systematic review presented in this chapter, I sought to quantify the additional evidential value of including preprints. This inclusion of preprint serves makes use of the preprint search tool presented in Chapter \@ref(sys-rev-tools-heading).

The results of this review are used to guide the primary analysis presented in Chapter \@ref(cprd-analysis-heading), in addition to forming a key evidence source used in the triangulation exercise presented in Chapter \@ref(discussion-heading).

@@

A pre-specified protocol for this analysis was registered using the Open Science Framework, and is available for inspection.[@mcguinnessluke2020] Deviations from this protocol are detailed in the relevant sections.

@@

In line with best-practice guidance, secondary reviewers were used to check the accuracy of screening, data extraction and risk-of-bias assessment processes. Due to the scale of the project, this systematic review was performed in conjunction with a team of secondary reviewers and an information specialist (see Acknowledgments and Author declaration). <!-- TODO add the following to acknowledgements These included Alexandra MacAleenan, Athena Sheppard, and Matthew Lee.  In addition, Sarah Dawson, an information specialist, provided input to the design of the search strategy.-->

@@

I systematically searched several electronic bibliographic databases to identify potentially relevant entries (hereafter referred to as "records"). The following databases were searched from inception onwards: Medline, EMBASE, Psychinfo, Cochrane Central Register of Controlled Trials (CENTRAL), and Web of Science Core Collection. As the contents of the Web of Science Core Collection can vary by institution,[@gusenbauer2020a] the specific databases searched via this platform are listed in Appendix \@ref(appendix-wos-databases). The search strategy used in each database was developed in an iterative manner using a combination of free text and controlled vocabulary (MeSH/EMTREE)[@lefebvre2019searching] terms to identify studies which have examined the relationship between blood lipids levels and dementia, incorporating input from an information specialist. The strategy included terms related to lipids, lipid modifying treatments, and dementia, and was designed for MEDLINE before being adapted for use in the other bibliography databases listed. An outline of the general strategy is presented in the Table \@ref(tab:searchOverview-table) below and the full search strategies for each database are presented in Appendix \@ref(appendix-search-strategy). <!-- TODO Need to actually attach each search strategy. Should be able to loop through search strategy results. -->

@@

When searching the bibliographic databases, study design filters were employed to try and reduce the screening load. To ensure that the study design filters are not excluding potentially relevant records, a random sample of 500 records identified by the main search but excluded by the filters (defined as "8 NOT 12" in Table \@ref(tab:searchOverview-table)) was screened.

@@

Records were imported into Endnote and de-duplicated using the method outlined in Bramer et al. (2016).[@bramer2016] In summary, this method uses multiple stages to identify potential duplicates, beginning with automatic deletion of records matching on multiple fields ("Author" + "Year" + "Title" + "Journal"), followed by manual review of less similar articles (e.g. those matched based on the "Title" field alone).

Following deduplication of records, screening (both title/abstract and full-text) was performed using a combination of Endnote, a citation management tool,[@hupe2019] and Rayyan, a web-based screening application.[@ouzzani2016] Title and abstract screening to remove obviously irrelevant records was performed primarily by me, with a random ~10% sample of excluded records being screened in duplicate to ensure consistency with the inclusion criteria. Additionally, records were rescreened by me with a 1 month lag to intra-rater consistency.

Similarly, I completed all full-text screening, with a random ~10% being screened in duplicate by a second reviewer. In addition, any records identified I identified as being difficult to assess against the inclusion criteria were screened in duplicate. Reasons for exclusion at this stage were recorded. Disagreements occurring during either stage of the screening process were resolved through discussion with a senior colleague. A PRIMSA flow diagram was produced to document how records moved through the review.[@page2021]

The criteria used to assess eligibility are presented in the subsequent sections.

@@

I sought to include studies that examine the relationship between blood lipid levels (or any specific lipid fraction, including total cholesterol, HDL, LDL, and triglycerides) and risk of incident dementia and its subtypes. Eligible study designs included randomized controlled trials and non-randomized observational studies of lipid modifying treatments, longitudinal studies examining the effect of increased/decreased blood lipid levels, and genetic instrumental variable (Mendelian randomization) studies examining the effect of genetically increased/decreased blood lipid levels.

Participants were screened for dementia at baseline and prevalent cases excluded. Alternatively, where no baseline screening was employed, participants were assumed to be dementia free if less than 50 years of age at baseline. Studies of any duration were included to allow for exploration of the effect of length of follow-up on the effect estimate using meta-regression. No limits were placed on the sample size of included studies.

Eligible studies defined dementia according to recognised criteria, for example the International Classification of Diseases (ICD),[@organizationwho1993] National Institute of Neurological Disorders and Stroke Association-Internationale pour la Recherche en l'Enseignement en Neurosciences (NINDS-AIREN),[@roman1993] or Diagnostic and Statistical Manual of Mental Disorders (DSM) criteria.[@edition2013] Studies utilising electronic health records were the exception to this, as it was assumed that these criteria were used when entering the outcome into the EHR.

@@

No limitations were imposed on publication status, publication date, venue or language, although sufficiently detailed reports of the studies to be able to examine their methods were required for inclusion.

@@

Due to the significant impact of a memory-related outcome such as dementia on exposure recall, case-control studies were excluded, though nested case-control studies, where historical records are used to determine the exposure status, were eligible for inclusion. Cross-sectional studies, qualitative studies, case reports/series and narrative reviews were also excluded. Studies which presented no evidence of attempting to exclude prevalent cases from their analyses were also excluded. Studies that measure change in continuous cognitive measures (e.g. MoCA score) without attempt to map these scores to ordinal groups (e.g. no dementia/dementia) were excluded. Conference abstracts with no corresponding full-text publication were examined, and where required, I contacted authors to obtain information on the study's status. Previous systematic reviews were not eligible for inclusion, but their reference lists were screened to identify any potentially relevant articles.

@@

Rater reliability was assessed using Gwet's agreement coefficient (AC1).[@gwet2008] This measure was chosen over other methods such as percent agreement (number of agreements divided by total number of assessments), as it accounts for chance agreement between reviewers but does not suffer from bias due to severely imbalanced marginal totals in the same way that Cohen's $kappa$ value does. [@cohen1960: @gwet2008; @wongpakaran2013] Given the small number of included studies in this review as a proportion of the total number screened, this is an important characteristic.

@@

#### Study-fication

As part of the data extraction process, multiple resords resulting from the analysis of the same data were included and grouped into single units, hereafter called studies. This is likely in the advent of multiple papers reporting results on the same cohort, but say, at different time points. Study-fication builds out the most comprehensive accounts of the studies and results from as many published articles were applicable.

@@

Following best practice, where summary data was presented across two groups (e.g. age at baseline stratified by hypercholesterolemia status), the following approach was used to combine the groups. [@higgins2019]

@@

This was implemented in a systematic manner, with the raw group data being extracted and a cleaning script written to combine the groups for analysis.

@@

The range of effect measures presented by studies (odds ratios, risk ratios, hazard ratios, etc) are not directly interchangeable in the context of systematic review. If the outcome is rare, as is the case for dementia outcomes at and estimated prevalence of  odds and risk ratios approximate each other.<!-- TODO CITATION NEEDED --> However, hazard ratios provide a very different interpretation, taking into account person-timea-at-risk in each treatment group. As such, different effect estimates can be one potential problem that precludes a meta-analysis of all studies.[@mckenzie2019]

Several existing reviews do not distinguish between the types of effect measures and include all existing studies in a single meta-analysis to produce an overall effect. In addition, there is some evidence of manipulation of effect estimates in previous reviews,(e.g. Chou, Sci Reports - at least one study disagrees with) but this is not accurately documented in the review text.
<!-- TODO Need to really fix this -->

In this review, studies reporting hazard ratios were synthesised separately to those reporting odds/risk ratios.

@@

A key use of the review presented in this chapter is to identify different sources of evidence at risk of a diverse range of biases, and to contrast and compare findings across them (see Section \@ref(triangulation-overview) for an overview of triangulation and Section \@ref(intro-triangulation) for the results of this qualitative analysis). To enable this triangulation exercise, a detailed and structured risk-of-bias assessment formed an important part of this review.

There has been a recent movement within the evidence synthesis community from examining _methodological quality_ to assessing _risk of bias_,[@mcguinness2018; @sterne2016] and thus directly evaluating the internal validity of a study. Internal validity is defined here as the absence of systematic error (or bias) in a study, which may influence its results.[@campbell1957; @juni2001]

This move was prompted by a unclear definition of "methodological quality" which could include facets such as unclear reporting, and challenges in the comparison of results from different tools. <!-- TODO CITATION NEEDED --> As part of this shift, the community also moved from checklist or score based tools towards domain-based methods, in which different potential sources of bias in a study are assessed in order. Additionally, bias should be assessed at the result (defined as a a specific outcome at a specific timepoint) rather than the study level. For example, a study may report on the efficacy of an intervention at six months and two years follow-up. In this case, missing outcome data that is not an issue at six months may introduce bias at 2 year follow-up, and assigning a bias judgement to the study as a whole masks the different level of bias for each unique result.

@@

Randomized controlled trials were assessed using the RoB2 tool.[@sterne2019] The tool assess the risk of bias across five domains: Bias arising from the randomization process, Bias due to deviations from intended intervention, Bias due to missing outcome data, Bias in measurement of the outcome, Bias in selection of the reported result. Acceptable judgements include: low risk of bias, some concerns, high risk of bias. Each of the 5 domains contains a series of signalling questions or prompts, which guide the user through the tool. Once a domain-level judgement for each domain has been assigned, an overall judgement, using the same three levels of risk of bias, is assigned to the result.

@@

For non-randomised studies of interventions (NRSI), I used the ROBINS-I (Risk Of Bias In Non-randomised Studies - of Interventions) tool.[@sterne2016] This tool assess the risk of bias across seven domains: Bias due to confounding, Bias due to selection of participants, Bias in classification of interventions, Bias due to deviations from intended interventions, Bias due to missing data, Bias in measurement of outcomes, and Bias in selection of the reported result. Similar to RoB 2, it has a number of prompting questions per domain, with acceptable judgements including “Low risk”, “Moderate risk”, “Serious risk” and “Critical risk”. Ideally, observational studies should be assessed in reference to an idealised randomised controlled trial. Under this approach, the (rare) overall judgement of "Low" indicates that the results should be considered equivalent to produced by a randomised controlled trial.

While a risk-of-bias tool for non-randomised studies of exposures (NRSE) is currently under development,[@morganr2020] but was insufficiently developed at the time the risk-of-bias assessments for this review were performed. Instead, I used a version of the ROBINS-I tool informed by the preliminary ROBINS-E tool (“Risk of Bias In Non-randomised Studies – of Exposure”), which I had applied in a published review.[@french2019] The version had no signalling questions and so judgements were made at the domain level. The motivation for this using this tool above other established tools such as the Newcastle-Ottowa scale (NOS).[@wells2000] was two-fold. In the first instance, as mentioned in the introduction to this section, using a domain-based tool has distinct advantages over better-developed checklist-type tools including the NOS. Additionally, using a domain-based tool for non-randomised studies of exposures enabled better comparison with risk-of-bias assessments performed for the other study designs.

@@

At present, no formalised risk-of-bias assessment tool for Mendelian randomization studies is available. Assessment of the risk of bias in Mendelian randomisation studies was informed by the approach used in a previous systematic review of Mendelian randomisation,[@mamluk2020] as identified by a review of risk-of-bias assessments in systematic reviews of Mendelian randomisation studies (advance results from this review were obtained from contact with the authors.). <!-- TODO CITATION NEEDED - add once published as a preprint --> A copy of this tool is available in Appendix \@ref(appendix-mr-rob), but in summary, results were assessed for bias arising from weak instruments, genetic and other confounding, pleiotropy, and selection of participants.

@@

A recent shift towards the assessment of missing evidence due to selective non-reporting - as distinct from the selective reporting of a single result from multiple planned - is demonstrated via the forthcoming RoB-ME (Risk of Bias due to Missing Evidence in a synthesis) tool.[@zotero-15123] <!-- TODO Need more here describing missing evidence --> The tool is in development stages, and as part of this review, I piloted the tool, and provided feedback to the developers. <!-- TODO Cite Ben's work on effect of missing study versus choice of analysis method -->

@@

An inital qualiatative synthesis of evidence was performed, summarising the data extracted from studies stratified by study design

Where individual studies were deemed comparable, they were incorporated into a quantitative analysis or "meta-analysis". <!-- TODO CITATION NEEDED --> Meta-analysis provides a summary or pooled effect estimate across studies.

Of note, studies were not combined across different study designs (i.e. RCTs were not combined in a meta-analysis with results from observational studies). The results from each individual analytical approach were summarised, but are compare and contrasted more fully in the triangulation exercise presented in Chapter \@ref(discussion-heading)

@@

@@

[Do I need to include]

@@

Random-effects methods estimate $\tau^2$ to estimate the common effect ($\mu$). There are many estimators of $\tau^2$ with varying assumptions [@Veroniki2016; @Bakbergenuly2020], the details of which are beyond the scope of this article. The estimate of $\tau^2$ is incorporated into the weighting of each study

@@

This was particularly important for the dose response meta-analysis, where the number of participants and the cut-offs per category were often not reported.

Several of the included studies presented data on multiple categories of lipid levels, but provided an overall effect estimate based on a comparison of only two of these categories (e.g. for example, highest vs lowest quartile). <!-- TODO Question: if I look at highest vs lowest, and also at top dose vs lowest dose, am I double counting results? --> While this allows for easy interpretation of the resulting effect estimate, it ignores any potential non-linear relationships between the exposure and outcome, in addition to discarding useful information contain in the interim groups. In order to address this limitation, I performed a dose-response meta-analysis in those studies reporting more than two categories for lipid levels.

Studies were excluded from this analysis if the number of categories was less than three, if the exposure cut-off points for for each category were not reported (e.g. if the study reports splitting participants into quartiles and comparing the highest vs lowest without giving the quartile bands).

A restricted cubic spline model was fitted to allow for a non-linear relationship, for example a U or J-shaped relationship, where low and high levels of the exposure can have different effects versus a "normal" reference dose. The locations of the knots in the model wer identified using fixed percentiles (25th, 50th, 75th) of the exposure data. Reference doses were defined _a priori_ as the cut-off of the "Normal"/"Optimal" categories for each fractions, as detailed in Table \@ref(tab:lipidLevels-table). Under this approach, the reference dose was defined as 200 mg/dL for total cholesterol, 100 mg/dL for LDL-c, 40 mg/dL for HDL-c, and 150 mg/dL for triglycerides.

Due to the requirements for the dose response analysis, studies were excluded from this secondary analysis if they did not provide the require information: cut-off points.

Where this was not reported in the study, I contacted the corresponding author to attempt to obtain the required information (see Section \@ref(contacting-authors)).

When the highest category was open ended (e.g LDL-c $\geqslant$ 200 mg/dL), I calculated category midpoint by assuming the width of the highest category was the same as the one immediately below it. Similarly, when, the lowest category was open-ended (e.g LDL-c $\leqslant$ 100 mg/dL), I set the lower boundary for this category to zero (though this is unlikely to occur natually, it was difficult to define).

@@

Where there was evidence of heterogeneity between results included in a meta-analysis, I investigated this further using meta-regression against reported characteristics. _A priori_, I was interested in the effect that the age at baseline, sex and risk-of-bias judgement had on the results. <!-- TODO Need more here -->

Finally, I investigated the potential for small study effects, which may be caused by publication bias, both visually using funnel plots and statistically using Egger's regression test.  <!-- TODO CITATION NEEDED - cite Jonathon on interpreting small study effects, and whatever citation is needed for  -->

@@

Evidence maps are useful way to explore the distribution of research cohorts included in a systematic review.[@saran2018] As such, the location of each individual study contributing to the evidence base was quantified and visualised on a world map.

>>>>

# Systematic review of all evidence available on the association between blood lipids and dementia outcomes {#sys-rev-heading}

@@

<!-- TODO Extra studies to include: -->
<!-- * Larsson - 10.1136/bmj.j5375 [@larsson2017b] -->

@@

Systematic reviews are a type of research study that aim to collect and combine all existing evidence to provide the best possible answer to an important research question. Well-performed reviews involve multiple steps including: searching of existing studies; assessment of the studies against predefined inclusion criteria; collection of data from each study; and assessment of each study's methods.

This chapter presents a systematic review of primary studies that have examined the relationship between the levels of blood lipids (such as cholesterol and triglycerides), and treatments that change these levels such as statins, and dementia outcomes.

My review included `r n_included` primary studies that contained information on this relationship. I found that statins appear to reduce the risk of Alzheimer's disease, but had no effect of vascular dementia. Lipid levels were not associated with any outcome. The methods used in some of the primary studies meant that I was less confident in the accuracy of their results.

@@

In this chapter, I describe a comprehensive systematic review of the relationship between blood lipid levels, and treatments that modify them, and the subsequent risk of dementia and related outcomes.

This analysis sought to address two specific aims. Firstly, as discussed in the Introduction to this thesis (Section \@ref(evidence-association)), several diverse forms of evidence on the relationship of lipids and dementia exist. These include randomised controlled trials, observational studies of different analytical design, and Mendelian randomisation studies. However, based on a scoping review of existing literature, no previous evidence synthesis exercise has attempted to examine the association of lipids/statins with dementia outcomes across these distinct evidence types. Collating these diverse evidence sources is important, as if the observed association between lipids and dementia is constant across them, it increases our confidence in the association. As such, the primary aim of this analysis was to systematically review all available literature describing prospective analyses, regardless of study design.

Secondly, I explicitly sought to include health-related preprint servers as a potential evidence source in this review, as they are infrequently considered by evidence synthesists but  report relevant unpublished analyses. As a sensitivity analysis to this review presented in this chapter, I sought to quantify the additional evidential value of including preprints, making use of the preprint search tool presented in Chapter \@ref(sys-rev-tools-heading).

The results of this review are used to guide the primary analysis presented in Chapter \@ref(cprd-analysis-heading) and \@ref(ipd-heading), in addition to forming a key evidence source used in the triangulation exercise presented in Chapter \@ref(tri-heading).

@@

A pre-specified protocol for this analysis was registered on the Open Science Framework platform and is available for inspection.[@mcguinnessluke2020] Deviations from this protocol are detailed in the relevant sections.

@@

In line with best-practice guidance, secondary reviewers were used to check the accuracy of screening, data extraction and risk-of-bias assessment processes. Due to the scale of the project, this review was performed in conjunction with a team of secondary reviewers (see Acknowledgements and Author Declaration in the front matter). <!-- TODO add the following to acknowledgements These included Alexandra MacAleenan, Athena Sheppard, and Matthew Lee.  In addition, Sarah Dawson, an information specialist, provided input to the design of the search strategy.-->

@@

I systematically searched several electronic bibliographic databases to identify potentially relevant entries (hereafter referred to as "records"). The following databases were searched from inception onwards: Medline, EMBASE, Psychinfo, Cochrane Central Register of Controlled Trials (CENTRAL), and Web of Science Core Collection. As the contents of the Web of Science Core Collection can vary by institution,[@gusenbauer2020a] the specific databases and date ranges for each database searched via this platform are listed in Appendix \@ref(appendix-wos-databases). The search strategy used in each database was developed in an iterative manner using a combination of free text and controlled vocabulary (MeSH/EMTREE)[@lefebvre2019searching] terms to identify studies which have examined the relationship between blood lipids levels and dementia, incorporating input from an information specialist. The strategy included terms related to lipids, lipid modifying treatments, and dementia, and was designed for MEDLINE before being adapted for use in the other bibliography databases listed. A high-level outline of the strategy is presented in the Table \@ref(tab:searchOverview-table) below and the full search strategies for each database are presented in Appendix \@ref(appendix-search-strategy). <!-- TODO Need to actually attach each search strategy. Should be able to loop through search strategy results. -->

@@

When searching the bibliographic databases, study design filters were employed to try and reduce the screening load. To ensure that the study design filters were not excluding potentially relevant records, a random sample of 500 records identified by the main search but excluded by the filters (defined as "8 NOT 12" in Table \@ref(tab:searchOverview-table)) was screened.

@@

Records were imported into Endnote and de-duplicated using the method outlined in Bramer et al. (2016).[@bramer2016] In summary, this method uses multiple stages to identify potential duplicates, beginning with automatic deletion of records matching on multiple fields ("Author" + "Year" + "Title" + "Journal"), followed by manual review of less similar articles (e.g. those identified as duplicates based on the "Title" field alone).

Following de-duplication of records, screening (both title/abstract and full-text) was performed using a combination of Endnote, a citation management tool,[@hupe2019] and Rayyan, a web-based screening application.[@ouzzani2016] Title and abstract screening to remove obviously irrelevant records was performed primarily by me, with a random ~10% sample of excluded records being screened in duplicate to ensure consistency with the inclusion criteria. Additionally, I re-screened the same ~10% sample with 3 month lag to assess intra-rater consistency.

Similarly, I completed all full-text screening, with a random ~20% being screened in duplicate by a second reviewer, in addition to any records identified I identified as being difficult to assess against the inclusion criteria were screened in duplicate. Reasons for exclusion at this stage were recorded. Disagreements occurring during either stage of the screening process were resolved through discussion with a senior colleague. A PRIMSA flow diagram was produced to document how records moved through the review.[@page2021]

The criteria against which records were assessed for eligibility are presented in the subsequent sections.

@@

I sought to include studies that examined blood lipid levels as a risk factor for demenita outcomes, defined either as binary hypercholesterolemia variable or by category/1-standard-deviation increase of a specific lipid fraction (total cholesterol, high- and low-density lipoprotein cholesterol and triglycerides). I also aimed to include studies examining the effect of treatments that modify lipids levels as a source of indirect evidence. Eligible study designs included randomized controlled trials and non-randomized observational studies of lipid modifying treatments, longitudinal studies examining the effect of increased/decreased blood lipid levels, and genetic instrumental variable (Mendelian randomization) studies examining the effect of genetically increased/decreased blood lipid levels.

Eligible studies screened participants for dementia at baseline and excluded any prevalent cases. Alternatively, where no baseline screening was employed, participants were assumed to be dementia free if less than 50 years of age at baseline. Studies of any duration were included to allow for exploration of the effect of length of follow-up on the effect estimate using meta-regression. No limits were placed on the sample size of included studies.

Eligible studies defined dementia outcomes according to recognised criteria, for example the International Classification of Diseases (ICD),[@organizationwho1993] National Institute of Neurological Disorders and Stroke Association-Internationale pour la Recherche en l'Enseignement en Neurosciences (NINDS-AIREN),[@roman1993] or Diagnostic and Statistical Manual of Mental Disorders (DSM) criteria.[@edition2013] Studies utilising electronic health records were the exception to this, as it was assumed that a valid criteria was employed when entering used when entering the outcome into the EHR.

Conference abstracts with no corresponding full-text publication were eligible, and where required, I contacted authors to obtain information on the study's status. No limitations were imposed on publication status, date, venue or language.

@@

@@

Due to the significant impact of a memory-related outcome such as dementia on exposure recall, case-control studies were excluded, though nested case-control studies, where historical records are used to determine the exposure status, were eligible for inclusion. Cross-sectional studies, qualitative studies, case reports/series and narrative reviews were also excluded, as were studies that measure change in continuous cognitive measures (e.g. MoCA score) without attempt to map these scores to ordinal groups (e.g. no dementia/dementia). Previous systematic reviews were not eligible for inclusion, but their reference lists were screened to identify any potentially relevant articles.

@@

Rater reliability was assessed using Gwet's agreement coefficient (AC1).[@gwet2008] This measure was chosen over other methods such as percent agreement (number of agreements divided by total number of assessments), as it accounts for chance agreement between reviewers but does not suffer from bias due to severely imbalanced marginal totals in the same way that Cohen's $kappa$ value does. [@cohen1960: @gwet2008; @wongpakaran2013] Given the small number of included studies in this review as a proportion of the total number screened, this is a useful characteristic.

@@

#### Grouping multiple reports into studies

As part of the data extraction process, multiple records resulting from the analysis of the same data were included and grouped into single units, hereafter called studies. This was common in cases where multiple papers report results on the same cohort but at different time points. This process builds out the most comprehensive account of a given studies possible by incorporating information from all available records.

@@

Following best practice, where summary data was presented across two groups (e.g. age at baseline stratified by hypercholesterolemia status), the following approach was used to combine the groups:[@higgins2019]

@@

This was implemented in a systematic manner, with the raw group data being extracted and a cleaning script employed to combine the groups for analysis.

@@

The range of effect measures presented by studies (odds ratios, risk ratios, hazard ratios, etc) are not directly interchangeable in the context of systematic review.  As such, different effect estimates can be one potential problem that precludes a meta-analysis of all studies.[@mckenzie2019] If the outcome is rare, as is the case for dementia outcomes, the estimated prevalence of  odds and risk ratios will approximate each other.<!-- TODO CITATION NEEDED --> However, hazard ratios provide a very different interpretation, taking into account person-time-at-risk in each treatment group.

Several existing reviews do not distinguish between the types of effect measures and include all existing studies in a single meta-analysis to produce an overall effect estimate. However, in this review, the small subset of studies reporting odds/risk ratios and synthesised separately to those reporting hazard ratios.

@@

A key aim of the review presented in this chapter is to identify different sources of evidence at risk of a diverse range of biases, and to contrast and compare findings across them (see Section \@ref(triangulation-overview) for an overview of triangulation and Chapter \@ref(tri-heading) for the results of this analysis). To enable this triangulation exercise, a detailed and structured risk-of-bias assessment formed an important part of this review.

There has been a recent movement within the evidence synthesis community away from examining _methodological quality_ to assessing _risk of bias_,[@mcguinness2018; @sterne2016] and thus directly evaluating the internal validity of a study. Internal validity is defined here as the absence of systematic error (or bias) in a study, which may influence its results.[@campbell1957; @juni2001]

This move was prompted by a unclear definition of "methodological quality" which could include facets such as unclear reporting, in addition to challenges in the comparison of results from different tools. <!-- TODO CITATION NEEDED --> As part of this shift, the focus shifted from checklist or score based tools towards domain-based methods, in which different potential sources of bias in a study are assessed in order. <!-- TODO CITATION NEEDED -->
Finally, tools movide from assessing bias at the study level to considering each indivudal result reported seperately. For example, a study may report on the efficacy of an intervention at six months and two years follow-up. In this case, missing outcome data that is not an issue at six months may introduce bias after two years of follow-up, and assigning a single risk-of-bias judgement to the study as a whole masks the different biases applicable to each unique result.

@@

Randomized controlled trials were assessed using the RoB 2 tool.[@sterne2019] The tool assess the risk of bias across five domains: bias arising from the randomization process, bias due to deviations from intended intervention, bias due to missing outcome data, bias in measurement of the outcome, bias in selection of the reported result. Acceptable judgements for each domain include: "low risk", "some concerns", "high risk". Each of the five domains contains a series of signalling questions or prompts which guide the user through the tool. Once a domain-level judgement for each domain has been assigned, an overall judgement, using the same three levels of risk of bias, is assigned to the result.

@@

For non-randomised studies of interventions (NRSI), I used the ROBINS-I (Risk Of Bias In Non-randomised Studies - of Interventions) tool.[@sterne2016] This tool assess the risk of bias across seven domains: bias due to confounding, bias due to selection of participants, bias in classification of interventions, bias due to deviations from intended interventions, bias due to missing data, bias in measurement of outcomes, and bias in selection of the reported result. Similar to RoB 2, it has a number of prompting questions per domain, with acceptable judgements including “low risk”, “moderate risk”, “serious risk” and “critical risk”. In the context of the tool, observational studies are assessed in reference to an idealised randomised controlled trial. Under this approach, the (rare) overall judgement of "Low" indicates that the results should be considered equivalent to produced by a randomised controlled trial.

While a risk-of-bias tool for non-randomised studies of exposures (NRSE) is currently under development,[@morganr2020] but was insufficiently developed at the time the risk-of-bias assessments for this review were performed. Instead, I used a version of the ROBINS-I tool informed by the preliminary ROBINS-E tool (Risk of Bias In Non-randomised Studies – of Exposure), which I had previously applied in a published review.[@french2019] The version had no signalling questions and so judgements, using the same four levels of bias as ROBINS-I, were made at the domain level. The motivation using this tool above other established tools such as the Newcastle-Ottowa scale (NOS)[@wells2000] was two-fold. In the first instance, as mentioned in the introduction to this section, using a domain-based tool has distinct advantages over better-developed checklist-type tools including the NOS. Additionally, using a domain-based tool for non-randomised studies of exposures enabled better comparison with risk-of-bias assessments performed for the other study designs as part of this review.

@@

At present, no formalised risk-of-bias assessment tool for Mendelian randomization studies is available. Assessment of the risk of bias in Mendelian randomisation studies was informed by the approach used in a previous systematic review of Mendelian randomisation,[@mamluk2020] as identified by a review of risk-of-bias assessments in systematic reviews of Mendelian randomisation studies (advance results from this review were obtained from contact with the review authors.). <!-- TODO CITATION NEEDED - add once published as a preprint --> A copy of this tool is available in Appendix \@ref(appendix-mr-rob), but in summary, results were assessed for bias arising from weak instruments, genetic and other confounding, pleiotropy and population stratification. Acceptable judgements for each fo the 5 domains in the tool included "low", "moderate" and "high" risk of bias.

@@

In addition to assessing the risk of bias within each result contributing to a synthesis, I also assessed risk of bias due to missing evidence at the analysis level. This assessment examines evidence missing due to selective non-reporting - as distinct from the selective reporting of a single result from multiple planned - and was performed using the forthcoming RoB-ME (Risk of Bias due to Missing Evidence in a synthesis) tool.[@zotero-15123] <!-- TODO Need more here describing missing evidence --> The tool is in development stages, and as part of this review, I piloted the tool, and provided feedback to the developers. <!-- TODO Cite Ben's work on effect of missing study versus choice of analysis method -->

@@

An inital qualiatative synthesis of evidence was performed, summarising the data extracted from studies stratified by study design. Where individual studies were deemed comparable, they were incorporated into a quantitative analysis or "meta-analysis". <!-- TODO CITATION NEEDED -->

Results were not combined across different study designs (i.e. RCTs were not combined in a meta-analysis with results from observational studies). The summary effect estimates produced by meta-analysis of individual study designs are discussed, but are compared and contrasted more fully as part of the triangulation exercise presented in Chapter \@ref(discussion-heading)

@@

For

@@

@@

<!-- Random-effects methods estimate $\tau^2$ to estimate the common effect ($\mu$). There are many estimators of $\tau^2$ with varying assumptions [@Veroniki2016; @Bakbergenuly2020], the details of which are beyond the scope of this article. The estimate of $\tau^2$ is incorporated into the weighting of each study -->

@@

Several of the included studies presented data on multiple categories of lipid levels, but provided an overall effect estimate based on a comparison of only two of these categories (e.g. for example, highest vs lowest quartile). <!-- TODO Question: if I look at highest vs lowest, and also at top dose vs lowest dose, am I double counting results? --> While this allows for easy interpretation of the resulting effect estimate, it ignores any potential non-linear relationships between the exposure and outcome, in addition to discarding useful information contain in the interim groups. In order to address this limitation, I performed a dose-response meta-analysis in those studies reporting more than two categories for lipid levels. This marks a departure from the published protocol, as I was unaware that such a large number of studies would report dementia risk across multiple lipid categories.

Studies were excluded from this analysis if the number of categories was less than three or if the necessary information for synthesis (cut-off points, number of participants and number of events per category) was not available. A restricted cubic spline model was fitted to allow for a non-linear relationship, for example a U or J-shaped relationship, where low and high levels of the exposure can have different effects versus a "normal" reference dose. <!-- TODO CITATION NEEDED --> The locations of the knots in the model wer identified using fixed percentiles (25th, 50th, 75th) of the exposure data. Reference doses were defined _a priori_ as the cut-off of the "Normal"/"Optimal" categories for each fraction, as detailed in Table \@ref(tab:lipidLevels-table). Under this approach, the reference dose was defined as 200 mg/dL for total cholesterol, 100 mg/dL for LDL-c, 40 mg/dL for HDL-c, and 150 mg/dL for triglycerides.

When the highest reported category was open ended (e.g LDL-c $\geqslant$ 200 mg/dL), I calculated the category midpoint by assuming the width of the highest category was the same as the one immediately below it. Similarly, when the lowest category was open-ended (e.g LDL-c $\leqslant$ 100 mg/dL), I set the lower boundary for this category to zero (though this is unlikely to occur naturally).

@@

Where there was evidence of heterogeneity between results included in a meta-analysis, I investigated this further using meta-regression against reported characteristics. _A priori_, I was interested in the effect that the age at baseline, sex and risk-of-bias judgement had on the results. Syntheses with greater than 10 studies were assessed for heterogeneity across these covariates. <!-- TODO Need more here -->

Finally, I investigated the potential for small study effects, which may be caused by publication bias, both visually using funnel plots and formally using Egger's regression test. <!-- TODO CITATION NEEDED - cite Jonathon on interpreting small study effects, and whatever citation is needed for  -->

@@

Evidence maps are useful way to explore the distribution of research cohorts included in a systematic review.[@saran2018] As part of the initial descriptive synthesis, the location of each individual study contributing to the evidence base was quantified and visualised on a world map.

One of the limitations of current risk-of-bias assessments in systematic reviews is that they are often divorced from the results to which they refer, and are infrequently incorporated into the analysis. <!-- TODO CITATION NEEDED --> In response to this criticism, I developed a new visualisation tool was designed to allow for the production of "paired" forest plots, as recommended by the ROB2 publication, where a risk-of-bias assessment is presented alongside it's corresponding numerical results.[@sterne2019] This tool was developed as an adjunct to this thesis to aid in creating standardised risk-of-bias figures,[@mcguinness2020robvisPaper] and the "paired" forest plot functionality grew out of a collaboration with other researchers to design a modular method for creating custom forest plots.[@zotero-14999] <!-- Apparently collaboration is a big thing they want to see --> A summary of this tool is contained in Appendix \@ref(appendix-robvis), and all forest plots presented in this Chapter were created using this tool.
